Algebraic File Synchronization: Adequacy and Completeness

### 2018-07-20 p2/conclusion.tex
Text review

We introduced relations and concepts that have proven particularly
useful, including 
independent commands, % (\cref{def_indep}), 
simple sequences, % (\cref{def_min_simp}),
and relationships between domains of sets of command sequences ($\worksmeqsign$)
that allowed us to formalize the correctness of reconciliation.
Among our partial results were interesting properties of these concepts,
for example,
that all legal permutations of simple sequences are equivalent (\cref{simple_reorder_equiv}),
that command sequences can be transformed into simple sequences using syntactical rules (\cref{can_simplify}),
or that the combination of independent sequences is defined wherever the components are defined (\cref{combine_independent_sequences}).
We believe these results will prove useful outside their current usage as well.

>>>>

We introduced relations and concepts that have proven particularly
useful, including 
independent commands, % (\cref{def_indep}), 
simple sequences, % (\cref{def_min_simp}),
and relationships between domains of sets of command sequences ($\worksmeqsign$)
that allowed us to formalize the correctness of reconciliation.
Among our partial results were interesting properties of these concepts,
for example,
that all permutations of simple sequences satisfying a partial order are equivalent (\cref{simple_reorder_equiv}),
that command sequences can be transformed into simple sequences using syntactical rules (\cref{can_simplify}),
or that the combination of independent sequences is defined wherever the components are defined (\cref{combine_independent_sequences}).
We believe these results will prove useful outside their current usage as well.

### 2018-07-20 p2/rec-intro.tex
Text review

We work based on the assumption that to achieve this, we need
to apply to $\FS_B$ the commands that have been applied to $\FS_A$, and \emph{vice versa}.
As some commands may have been applied to both filesystems, our first approximation
is $\reca = \amb$ and $\recb = \bma$.
This, however, will break both filesystems if there have been incompatible updates
in $A$ and $B$. 
Our aim is therefore to provide an algorithm that selects the commands 
$\reca \subset A\setminus B$
and $\recb \subset B\setminus A$ 
so that $\reca\aFS_B\neq\fsbroken$ and $\recb\aFS_A\neq\fsbroken$,
and show that these are the longest sequences with this property, that is,
adding any further command to $\reca$ or $\recb$ will break the filesystems.

We assume that $A$ and $B$ are returned by an update detector,
and so they are simple sequences.
(Otherwise, they can be simplified to become simple sequences; see \cref{can_simplify}.)
We also assume that $\FS$, $\FS_A$ and $\FS_B$ are not broken.
With these assumptions,
we can now define our reconciliation algorithm.

@@

Note that $\alpha\in\reca$ is not an assertion command, and $\bmanp$ contains no assertion commands,
and $\alpha\not\in\bmanp$, therefore $\alpha\indep\bmanp$ is equivalent to
the node $\alpha$ affects being incomparable to the nodes commands in $\bmanp$ affect.
Therefore whether two commands are independent is easy to determine programmatically
in at most $O(x^2)$ time, where $x$ is the number of commands. % TODO
Then, to generate the sequence $\reca$ from the resulting set, 
apply topological sorting according to $\orderrel$ as defined in \cref{ordering}.

\myskip
Next, we would now like to prove that $\reca$ and $\recb$ can be applied to the replicas,
that is, without loss of generality,
\[ \reca\aFS_B\neq\fsbroken. \]
So that we can formalize statements needed to prove this,
let us introduce partial orders that describe under what conditions
sequences of commands work, that is, do not break a filesystem.

>>>>

We work based on the assumption that to achieve this, we need
to apply to $\FS_B$ the commands that have been applied to $\FS_A$, and \emph{vice versa}.
As some commands may have been applied to both filesystems, our first approximation
is to apply the commands in $\bmanp$ to $\FS_A$ and those in $\ambnp$ to $\FS_B$.
This, however, will break both filesystems if there have been incompatible updates
in $A$ and $B$. 
Our aim is therefore to provide an algorithm that selects the commands 
$\reca \subset A\setminus B$
and $\recb \subset B\setminus A$ 
so that $\reca\aFS_B\neq\fsbroken$ and $\recb\aFS_A\neq\fsbroken$,
and show that these are maximal in the sense that
adding any further command to $\reca$ or $\recb$ will break the filesystems.

We will work based on update sequences $A$ and $B$ that are simple.
If the reconciliation algorithm is to work on arbitrary sequences,
\cref{can_simplify} can be used to convert those sequences to simple ones.

@@

Note that $\alpha\in\reca$ is not an assertion command, and $\bmanp$ contains no assertion commands,
and $\alpha\not\in\bmanp$, therefore $\alpha\indep\bmanp$ is equivalent to
the node of $\alpha$ being incomparable to the nodes of the commands in $\bmanp$.
Therefore whether two commands are independent 
for the purposes of reconciliation
is easy to determine programmatically.
Then, to generate the sequence $\reca$ from the resulting set, 
apply topological sorting according to $\orderrel$ as defined in \cref{ordering}.

\myskip
We aim prove that $\reca$ and $\recb$ can be applied to the replicas,
that is, without loss of generality,
\[ \reca\aFS_B\neq\fsbroken. \]
So that we can formalize statements needed to prove this,
we introduce the partial order $\worksmeqsign$ that describes under what conditions
sequences of commands work, that is, do not break a filesystem.

### 2018-07-20 p2/rec-sound.tex
Added conclusion
Text review

We show that $A$ can be transformed into $A'$ for which this is true.
We know that if the marked commands are not at the beginning, then
the sequence contains an unmarked command followed by a marked command.
We show that these can be swapped resulting in an equivalent sequence.
Then, by repeating this process similarly to bubble sorting, we arrive at 
a suitable permutation of $A$, which is also equivalent to $A$, and therefore in $\orderset{A}$.

@@

We would like to prove that $\worksmbx{A,B}{B\cc \reca}$.
From \cref{can_move_intersection} above we know that we can move commands in $\acbnp$
to the beginning of $A$ and $B$, and so this is equivalent to
\[ \worksm{\big\{\acb\cc\amb,~ \acb\cc\bma\big\}}{\acb\cc \bma\cc \reca}. \]

For ease of reading, let us rename
\begin{itemize}
\item $\amb$ to $S$,
\item $\reca$ to $S^*$,
\item $\bma$ to $T$,
\item and $\acb$ to $U$.
\end{itemize}
In this notation, we intend to prove that
\[ \worksmbx{U\cc S,U\cc T}{U\cc T\cc S^*}. \]
We can do so if we can show that
\[ \worksmbx{S,T}{S^*}. \]

@@

To restate some results above, we therefore already know the following:
\begin{itemize}
\item $S=\amb$ and $T=\bma$ are simple as $A$ and $B$ are simple
\item $S\cap T = \amb \cap \bma = \emptyset$
\item $S^*$ is the largest subset of $S$ for which $S^*\indep T$, as
$\reca$ is the largest subset of $\amb$ for which $\reca\indep\bma$.
\end{itemize}
Our theorem is therefore proven if we prove the following lemma.
\newcommand{\condSimple}{(c1)}
\newcommand{\condDisj}{(c2)}
\newcommand{\condApr}{(c3)}
\begin{mylem}\mlabel{reconciliation_correct_part}
If
{\rm\condSimple} $S$ and $T$ are simple sequences,
{\rm\condDisj} $S\cap T=\emptyset$,
and {\rm\condApr} $S^*$ is the largest subset of $S$ where $S^*\indep T$,
then
$\worksmbx{S,T}{S^*}$.
\end{mylem}
\begin{proof}
This is trivial unless $\worksmnilb{S,T}$, so we assume that it is the case.
The proof is similar to that of \cref{can_move_intersection}.
We mark all commands in $S$ that are in its subset $S^*$, and
we prove that $S$ can be transformed into $S'$
where all marked commands are at the beginning.
If so, then 
$\worksm{S}{S^*}$ based on \cref{worksextpostfix},
from which we get $\worksmbx{S,T}{S^*}$.

>>>>

We know that if the marked commands are not at the beginning, then
the sequence contains an unmarked command followed by a marked command.
We show that these can be swapped resulting in an equivalent sequence.
Then, by repeating this process similarly to bubble sorting, we arrive at 
a suitable permutation of $A$, which is also equivalent to $A$, and therefore is in $\orderset{A}$.

@@

We aim to prove that $\worksmbx{A,B}{B\cc \reca}$.
From \cref{can_move_intersection} above we know that we can move commands in $\acbnp$
to the beginning of $A$ and $B$, and so this claim is equivalent to
\[ \worksm{\big\{\acb\cc\amb,~ \acb\cc\bma\big\}}{\acb\cc \bma\cc \reca}. \]

For ease of reading, let us rename
 $\amb$ to $S$,
 $\reca$ to $S^*$,
 $\bma$ to $T$,
 and $\acb$ to $U$.
In this notation, we intend to prove that
\[ \worksmbx{U\cc S,U\cc T}{U\cc T\cc S^*}. \]
We can do so if we can show that
\[ \worksmbx{S,T}{S^*}. \]

@@

To restate some results above, we therefore already know the following:
\begin{itemize}
\item $S=\amb$ and $T=\bma$ are simple as $A$ and $B$ are simple;
\item $S\cap T = \amb \cap \bma = \emptyset$;
\item $S^*$ is the largest subset of $S$ for which $S^*\indep T$, as
$\reca$ is the largest subset of $\amb$ for which $\reca\indep\bma$.
\end{itemize}
Our theorem is therefore proven if we prove that these imply the proposition above:
\newcommand{\condSimple}{(c1)}
\newcommand{\condDisj}{(c2)}
\newcommand{\condApr}{(c3)}
\begin{mylem}\mlabel{reconciliation_correct_part}
If
{\rm\condSimple} $S$ and $T$ are simple sequences,
{\rm\condDisj} $S\cap T=\emptyset$,
and {\rm\condApr} $S^*$ is the largest subset of $S$ where $S^*\indep T$,
then
$\worksmbx{S,T}{S^*}$.
\end{mylem}
\begin{proof}
This is trivial unless $\worksmnilb{S,T}$, so we assume that it is the case.
The proof is similar to that of \cref{can_move_intersection}.
We mark all commands in $S$ that are in its subset $S^*$, and
we prove that $S$ can be transformed into $S'$
where all marked commands are at the beginning.
If so, then 
$\worksm{S}{S^*}$ based on \cref{worksextpostfix},
from which we get $\worksmbx{S,T}{S^*}$.

### 2018-07-20 p2/works-inverse.tex
Resolve TODOs
Spellcheck
Text review

\begin{mydef}[Sets of sequences, their domains]
As we will frequently refer to sets of fequences,
we will use calligraphic letters (e.g. $\sqs{A}$, $\sqs{B}$, $\sqs{C}$ and $\sqs{S}$)
to denote such sets for brevity.
We will write $\Dom{\sqs{A}}$ for $\bigcap_{A\in\sqs{A}} \Dom{A}$,
and $\sqs{A}\cc\sqs{B}$ for $\{A\cc B\whr A\in\sqs{A}, B\in\sqs{B}\}$.
\end{mydef}

@@

\begin{mydef}[$\worksmeqsign$]\mlabel{def_works}
For two sets of sequences $\sqs{A}$ and $\sqs{B}$
we write $\worksm{\sqs{A}}{\sqs{B}}$ iff $\Dom{\sqs{A}} \subseteq \Dom{\sqs{B}}$;
that is, iff all sequences in $\sqs{B}$ are defined on (do not break)
all filesystems on which sequences in $\sqs{A}$ are defined.
\end{mydef}

When $\sqs{A}$ or $\sqs{B}$ contains a single sequence,
we leave out the curly brackets and write,
e.g. $A\cc\sqs{B}$ to mean $\{A\}\cc\sqs{B}$,
or $\worksm{A}{B}$ to mean $\worksmbb{A}{B}$.
Also, we write $\worksmnil{\sqs{A}}$ to mean that
there are filesystems on which all sequences in $\sqs{A}$ are defined.
If $\sqs{A}$ contains a single sequence only, $A$, this is equivalent to $\wrks{A}$.

@@

\begin{myclm}\mlabel{works_restricted}
$\forall \sqs{A},\sqs{B},S: \worksm{\sqs{A}}{\sqs{B}} \Rightarrow \worksm{S\cc \sqs{A}}{S\cc \sqs{B}}$,
that is,
the relationship between the domains of the sets $\sqs{A}$ and $\sqs{B}$ 
does not change if both are prefixed by a sequence $S$.
\end{myclm}
\begin{proof}
This is because the function $S$ is a binary relation between filesystems,
and we can treat its inverse relation as a one-to-many mapping between filesystems
that maps $\Dom{\sqs{A}}$ to $\Dom{S\cc\sqs{A}}$ and $\Dom{\sqs{B}}$ to $\Dom{S\cc\sqs{B}}$.
As such a one-to-many mapping maps a subset of a set to a subset of the image of the set,
if $\Dom{\sqs{A}}\subseteq\Dom{\sqs{B}}$, then $\Dom{S\cc\sqs{A}}\subseteq\Dom{S\cc\sqs{B}}$.
\end{proof}

@@

\begin{mylem}\mlabel{combine_independent_commands}
The combination of independent commands is defined on all filesystems
where both of the commands are defined:
\[ \alpha\indep \beta \Rightarrow \worksmbx{\alpha, \beta}{\alpha\cc \beta}. \]
\end{mylem}
\begin{proof}
% In the following proof we reach back to our filesystem model.
The proposition follows from \cref{independent_details},
and we follow the three cases listed there.

The first case is that $n\unrel m$ where $\alpha=\cxynv$, $\beta=\czwmv$.
Using an inverse proof we assume that for a filesystem $\FS$
both $\cxynv\aFS$ and $\czwmv\aFS$ are defined, but $(\cxynv\cc\czwmv)\FS$ is broken.
Since $n\neq m$, we know the input types of the command are compatible with $\FS$,
so $\FS$ must be broken because it violates the tree property.

@@

\begin{mylem}\mlabel{combine_independent_sequences}
The combination of independent sequences is defined on all filesystems
where both of the sequences are defined:
\[ S\indep T \Longrightarrow \worksmbx{S,T}{S\cc T}. \]
\end{mylem}
\begin{proof}
Assume that there is a filesystem $\FS$ so that
$S\aFS\neq\fsbroken$ and $T\aFS\neq\fsbroken$, but
$(S\cc T)\FS=\fsbroken$.

From \cref{def_indep} we know that
the commands in $S$ and $T$ pairwise commute, and so any sequence
that contains the commands from $S$ and $T$ and preserve their original partial order
is equivalent to $S\cc T$ on all filesystems.

@@

\begin{mylem}\mlabel{worksinputmatch}
If $S$ and $T$ are minimal sequences, $\worksmnilb{S,T}$,
and there are commands $\cxynv\in S$ and $\czwnv\in T$ on the same node $n$,
then the input types of these commands must match, i.e. $[x]=[z]$.
\end{mylem}
\begin{proof}
This result is similar to \cref{equiv_simple_same_commands} and
is easily shown using a proof by contradiction: if $[x]\neq [z]$, then there is no filesystem that
either $\cxynv$ or $\czwnv$ would not break, 
and consequently $S$ and $T$ cannot work on the same filesystem.
\end{proof}

>>>>

\begin{mydef}[Sets of sequences, their domains]
As we will frequently refer to sets of sequences,
we will use calligraphic letters (e.g. $\sqs{A}$, $\sqs{B}$, $\sqs{C}$ and $\sqs{S}$)
to denote such sets for brevity.
We will write $\Dom{\sqs{A}}$ for $\bigcap_{A\in\sqs{A}} \Dom{A}$,
and $\sqs{A}\cc\sqs{B}$ for $\{A\cc B\whr A\in\sqs{A}, B\in\sqs{B}\}$.
\end{mydef}

@@

\begin{mydef}[$\worksmeqsign$]\mlabel{def_works}
For two sets of sequences $\sqs{A}$ and $\sqs{B}$
we write $\worksm{\sqs{A}}{\sqs{B}}$ iff $\Dom{\sqs{A}} \subseteq \Dom{\sqs{B}}$;
that is, iff all sequences in $\sqs{B}$ are defined on (do not break)
all filesystems on which sequences in $\sqs{A}$ are defined.
We write $\sqs{A}\worksmsign\sqs{B}$ iff $\Dom{\sqs{A}} \subset \Dom{\sqs{B}}$.
\end{mydef}

When $\sqs{A}$ or $\sqs{B}$ contains a single sequence,
we leave out the curly brackets and write,
e.g. $A\cc\sqs{B}$ to mean $\{A\}\cc\sqs{B}$,
or $\worksm{A}{B}$ to mean $\worksmbb{A}{B}$.
Also, we write $\worksmnil{\sqs{A}}$ to mean that
there is at least one filesystem on which all sequences in $\sqs{A}$ are defined.
If $\sqs{A}$ contains a single sequence only, $A$, this is equivalent to $\wrks{A}$.

@@

\begin{myclm}\mlabel{works_restricted}
$\forall \sqs{A},\sqs{B},S: \worksm{\sqs{A}}{\sqs{B}} \Rightarrow \worksm{S\cc \sqs{A}}{S\cc \sqs{B}}$,
that is,
the relationship between the domains of the sets $\sqs{A}$ and $\sqs{B}$ 
does not change if both are prefixed by a sequence $S$.
\end{myclm}
\begin{proof}
This is because the sequence $S$, as an endofunction on filesystems, is a binary relation,
and we can treat its inverse relation as a one-to-many mapping between filesystems
that maps $\Dom{\sqs{A}}$ to $\Dom{S\cc\sqs{A}}$ and $\Dom{\sqs{B}}$ to $\Dom{S\cc\sqs{B}}$.
As such a one-to-many mapping maps a subset of a set to a subset of the image of the set,
if $\Dom{\sqs{A}}\subseteq\Dom{\sqs{B}}$, then $\Dom{S\cc\sqs{A}}\subseteq\Dom{S\cc\sqs{B}}$.
\end{proof}

@@

\begin{mylem}\mlabel{combine_independent_commands}
The combination of independent commands is defined on all filesystems
where both of the commands are defined:
\[ \alpha\indep \beta \Rightarrow \worksmbx{\alpha, \beta}{\alpha\cc \beta}. \]
\end{mylem}
\begin{proof}
% BACK_TO_FS
% In the following proof we reach back to our filesystem model.
The proposition uses \cref{independent_details},
and we follow the three cases listed there.
Let $\alpha=\cxynv$ and $\beta=\czwmv$.

The first case is that $n\unrel m$ where $\alpha=\cxynv$, $\beta=\czwmv$.
Assume, by contradiction, that for a filesystem $\FS$
both $\cxynv\aFS$ and $\czwmv\aFS$ are defined, but $(\cxynv\cc\czwmv)\aFS$ is broken.
Since $n\neq m$, we know the input types of the command are compatible with $\FS$,
so $(\cxynv\cc\czwmv)\aFS$ must be broken because it violates the tree property.

@@

\begin{mylem}\mlabel{combine_independent_sequences}
The combination of independent sequences is defined on all filesystems
where both of the sequences are defined:
\[ S\indep T \Longrightarrow \worksmbx{S,T}{S\cc T}. \]
\end{mylem}
\begin{proof}
Assume, by contradiction, that there is a filesystem $\FS$ so that
$S\aFS\neq\fsbroken$ and $T\aFS\neq\fsbroken$, but
$(S\cc T)\FS=\fsbroken$.

From \cref{def_indep}
the commands in $S$ and $T$ pairwise commute, and so any sequence
that contains the commands from $S$ and $T$ and preserve their original partial order
is equivalent to $S\cc T$ on all filesystems.

@@

\begin{mylem}\mlabel{worksinputmatch}
If $S$ and $T$ are minimal sequences, $\worksmnilb{S,T}$,
and there are commands $\cxynv\in S$ and $\czwnv\in T$ on the same node $n$,
then the input types of these commands must match, i.e. $[x]=[z]$.
\end{mylem}
\begin{proof}
This result is similar to \cref{equiv_simple_same_commands}.
If, by contradiction, $[x]\neq [z]$, then there would be no filesystem that
either $\cxynv$ or $\czwnv$ would not break, 
and consequently $S$ and $T$ could not work on the same filesystem.
\end{proof}
